---
title: "Examining predictive pricing indicators for cryptocurrency assets"
author: "Tiffany Valdecantos, Terrill Toe, Adetayo Adegoke"
date: "`r format(Sys.Date())`"
csl: vancouver.csl
bibliography: references.bib
font-size: 11pt
link-citations: true
output: 
  pdf_document:
    toc: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
#install.packages("readr")
install.packages("corrplot")                 # Install reshape2 package
suppressMessages(library("reshape2"))
suppressMessages(library(readr))
suppressMessages(library(caret))
suppressMessages(library(UsingR))
suppressMessages(library(quantmod))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(tseries))
suppressMessages(library(gridExtra))
suppressMessages(library(GGally))
suppressMessages(library(corrplot))
suppressMessages(library(pROC))
suppressMessages(library(MLmetrics))
# Source our R script with functions
source("src/quant_funcs.R")
```

# Research Question and Motivation

According to the US government's Internal Revenue Service, a virtual currency is a digital representation of value, other than a representation of the U.S. dollar or a foreign currency ("real currency"), that functions as a unit of account, a store of value, and a medium of exchange. The agency also defines cryptocurrencies as a type of virtual currency that uses cryptography to secure transactions that are digitally recorded on a distributed ledger, such as a blockchain @CryptoDefIrs.

Cryptocurrencies are mainstream assets that are electronically traded, similar to stocks and other equity types. There are many different cryptocurrencies, with several more being introduced to the public every few months for the opportunity to make transactions based on these assets. There is value to examining factors that drive cryptocurrency asset prices, since the market size of these currencies is currently over 1 trillion USD in traditional fiat USD, and is projected to grow to over 32 trillion USD by 2027 @CryptoMarketSizeYahoo. Because of their growing market size, cryptocurrencies have a significant influence on the global economy, and understanding cryptocurrency asset class pricing predictors can inform governments, businesses and retail investors.

This research paper examines the following questions as it pertains to cryptocurrencies:

1.  Is the fed interest rate correlated to the price of Bitcoin?
2.  Is the fed unemployment rate correlated to the price of Bitcoin?
3.  What is the relationship between the price of Solarium and Ethereum? Are they correlated?
4.  What is the relationship between the price of Bitcoin and USD coin? Are they correlated?
5.  Can an effective model be developed to predict the price of Bitcoin based on the predictors examined?

# Hypothesis

# Data Engineering & Visualization

```{r,eval=TRUE}

# load cryptocurrency data
# sol_data <- read_csv("solana.csv")
#sol2_data <- read_csv("solana2.csv", show_col_types = FALSE)
btc_data <- read.csv("archive/coin_Bitcoin.csv")
#eth_data <- read.csv("archive/coin_Ethereum.csv")
# usdc_data <- read_csv("usdcoin.csv")

#set.seed(25)
# str(blockchain_data)

# create solana (sol) cryptocurrency test and training data sets
# sol_ind <- createDataPartition(sol_data$Close, p = 0.6, list = F)
# sol_training <- sol_data[sol_ind,] # training data set to help develop model
# sol_testing <- sol_data[-sol_ind,] # testing data set to help validate model - predicted vs observed outcomes

# create solana (sol) cryptocurrency test and training data sets
#sol2_ind <- createDataPartition(sol2_data$Close, p = 0.6, list = F)
#sol2_training <- sol2_data[sol2_ind,] # training data set to help develop model
#sol2_testing <- sol2_data[-sol2_ind,] # testing data set to help validate model - predicted vs observed outcomes

# # create bitcoin (btc) cryptocurrency test and training data sets
# btc_ind <- createDataPartition(btc_data$Close, p = 0.6, list = F)
# btc_training <- btc_data[btc_ind,]
# btc_testing <- btc_data[-btc_ind,]
# 
# # create ethereum (eth) cryptocurrency test and training data sets
# eth_ind <- createDataPartition(eth_data$Close, p = 0.6, list = F)
# eth_training <- eth_data[eth_ind,]
# eth_testing <- eth_data[-eth_ind,]
# 
# # create usd coin (usdc) cryptocurrency test and training data sets
# usdc_ind <- createDataPartition(usdc_data$Close, p = 0.6, list = F)
# usdc_training <- usdc_data[usdc_ind,]
# usdc_testing <- usdc_data[-usdc_ind,]
# 
# # sapply(lapply(sol_training, unique), length)
# sol_cmplx_model <- lm(Close ~ . -Name -Symbol, data = sol_training)
# summary(sol_cmplx_model)

#sol2_cmplx_model <- lm(Close ~ ., data = sol2_training)
#summary(sol2_cmplx_model)

#stepAIC(sol2_cmplx_model, trace = F)
#sol2_optimized_model <- lm(Close ~ TotalPublicDebt + SP500Close, data = sol2_training)
#summary(sol2_optimized_model)

#sol3 <- lm(Close ~ SP500Open, data = sol2_training)
#summary(sol3)
str(btc_data)
```

Data munging steps to do:
1. convert Date to Date type
2. Transform Volumn and Maeketcap so it is within the scale of the remaining data
3. Check for any null values within the original data
```{r}
#Convert the Date column to Date and the 
btc_data$Date <- as.Date(btc_data$Date, format = "%Y-%m-%d")
#Transform the Volume and Marketcap columns to scale the data
btc_data$Marketcap <- btc_data$Marketcap / 1000000000000
btc_data$Volume <- btc_data$Volume / 100000000000
# Feature engineering of some new columns
# Sort the data by date in decendeing order
btc_data <- btc_data[order(btc_data$Date),]
#Check for any null values in the data set
btc_data %>% summarise(across(everything(), ~ sum(is.na(.))))
```
Now let's do some feature engineering by adding some columns that would give us future and past results.


Feature Definition:

Bit_returns -> the simple daily returns using the closing price at time t and t-1
Bit_log_returns - the daily log using the closing price
Bit_days_range - the range in price for that given date, highest price - the lowest price of that day.
Bit_log_wtd_ret - the week to date returns using the closing price at time t and t-7
Bit_log_biwtd_ret - the bi weekly to date returns using the closing price at time t and t-14
Bit_log_mtd_ret - the month to date returns using the closing price at time t and t-30
ema_log_ret_7 - the exponential moving average of the week to date returns. Weighted average using values in range of time t and t-7
ema_log_ret_14 - the exponential moving average of the bi weekly to date returns. Weighted average using values in range of time t and t-14
ema_log_ret_30 - the exponential moving average of the month to date returns. Weighted average using values in range of time t and t-30
moving_avg_10 - the 10 moving average of the closing price
moving_avg_20 - the 20 moving average of the closing price
moving_avg_50 - the 50 moving average of the closing price
moving_avg_50_diff - the differnece between the 50 day moving average at t and t-1
moving_avg_50_diff_7 - the differnece between the 50 day moving average at t and t-7
Bit_rsi_3 - Momentum indicator using 30 days of prior prices at time t
Bit_rsi_60 - Momentum indicator using 60 days of prior prices at time t
Bit_rsi_90 - Momentum indicator using 90 days of prior prices at time t

resources:
RSI indicator: https://www.investopedia.com/terms/r/rsi.asp
differencing in Time series: 

Stationarity and Differencing: https://towardsdatascience.com/an-intuitive-guide-to-differencing-time-series-in-python-1d6c7a2c067a

Exponential Moving Average (EMA): https://www.investopedia.com/terms/e/ema.asp
```{r}
# Get the daily log returns
btc_data$Bit_log_returns <- c(NA, diff(log(btc_data$Close), lag=1))
btc_data$Bit_log_returns <- round(btc_data$Bit_log_returns, 5)
# Get the daily returns
n <- nrow(btc_data)
prices <- btc_data$Close
returns<- ((prices[2:n] - prices[1:(n-1)])/prices[1:(n-1)])
btc_data$Bit_returns <- c(NA, returns)
btc_data$Bit_returns <- round(btc_data$Bit_returns, 5)
# Get the daily range
btc_data$Bit_days_range <- btc_data$High - btc_data$Low
# Get the week(7 days) to date returns
btc_data$Bit_log_wtd_ret <- c(rep(NA,7), diff(log(btc_data$Close), lag=7))
# Get the week(7 days) to date returns - the past 7 days leading up to the current records time period
btc_data$Bit_log_biwtd_ret <- c(rep(NA,14), diff(log(btc_data$Close), lag=14))
# Get the Bi-weekly(14 days) to date returns
btc_data$Bit_log_mtd_ret <- c(rep(NA,30), diff(log(btc_data$Close), lag=30))
#Get the 7 day exponential moving average of daily returns
btc_data$ema_log_ret_7 <- EMA(btc_data$Bit_log_returns, 7)
#Get the 14 day exponential moving average of daily returns
btc_data$ema_log_ret_14 <- EMA(btc_data$Bit_log_returns, 14)
#Get the 30 day exponential moving average of daily returns
btc_data$ema_log_ret_30 <- EMA(btc_data$Bit_log_returns, 30)
#Create variables for moving averages
btc_data <- mutate(btc_data, moving_avg_10 = rollmean(Close, 10, fill = 0, align = "right"))
btc_data <- mutate(btc_data, moving_avg_20 = rollmean(Close, 20, fill = 0, align = "right"))
btc_data <- mutate(btc_data, moving_avg_50 = rollmean(Close, 50, fill = 0, align = "right"))
#Get the difference between the moving averages
btc_data$moving_avg_50_diff <- c(rep(NA, 2), diff(btc_data$moving_avg_50, lag = 1, differences = 2))
btc_data$moving_avg_50_diff_7 <- c(rep(NA, 14), diff(btc_data$moving_avg_50, lag = 7, differences = 2))
# Get the RSI indicator, for a monthly period, 60 day period and a 90 day period
btc_data$Bit_rsi_3 <- RSI(btc_data$Close, n=30, maType = "EMA") / 100
btc_data$Bit_rsi_60 <- RSI(btc_data$Close, n=60, maType = "EMA") / 100
btc_data$Bit_rsi_90 <- RSI(btc_data$Close, n=90, maType = "EMA") / 100

summary(btc_data)
```

Because we are dealing with time series data, we will run the Augmented Dickey fuller test on our continuous variables to test for stationarity.

Augmented dickey fuller test: https://www.statology.org/dickey-fuller-test-in-r/
```{r}

print("################## Testing for the Close ##################")
adf.test(btc_data$Close)

print("################## Testing for the log Returns ##################")
adf.test(btc_data$Bit_log_returns[2:n])

print("################## Testing for the Simple Returns ##################")
adf.test(btc_data$Bit_returns[2:n])

print("################## Testing for the Volume ##################")
adf.test(btc_data$Volume)

print("################## Testing for the Marketcap ##################")
adf.test(btc_data$Marketcap)

print("################## Testing for the Day's Range ##################")
adf.test(btc_data$Bit_days_range)

print("################## Testing for the WTD Returns ##################")
adf.test(btc_data$Bit_log_wtd_ret[8:n])

print("################## Testing for the Bi -WTD Returns ##################")
adf.test(btc_data$Bit_log_biwtd_ret[15:n])

print("################## Testing for the MTD Returns ##################")
adf.test(btc_data$Bit_log_mtd_ret[31:n])

print("################## Testing for the 7 day EMA Daily Returns ##################")
adf.test(btc_data$ema_log_ret_7[8:n])

print("################## Testing for the 14 day EMA Daily Returns ##################")
adf.test(btc_data$ema_log_ret_14[15:n])

print("################## Testing for the 30 day EMA Daily Returns ##################")
adf.test(btc_data$ema_log_ret_30[31:n])

print("################## Testing for 30 day RSI ##################")
adf.test(btc_data$Bit_rsi_3[31:n])

print("################## Testing for 60 day RSI ##################")
adf.test(btc_data$Bit_rsi_60[61:n])

print("################## Testing for 90 day RSI ##################")
adf.test(btc_data$Bit_rsi_90[91:n])

print("################## Testing for 10 day moving average ##################")
adf.test(btc_data$moving_avg_10[11:n])

print("################## Testing for 20 day moving average ##################")
adf.test(btc_data$moving_avg_20[21:n])

print("################## Testing for 50 day moving average ##################")
adf.test(btc_data$moving_avg_50[21:n])

print("################## Testing for the 1 Day Difference between 50 day moving average ##################")
adf.test(btc_data$moving_avg_50_diff[100:n])

print("################## Testing for the 7 Day Difference between 50 day moving average ##################")
adf.test(btc_data$moving_avg_50_diff_7[110:n])


```

From the Dickey-Fuller Test we can see that from the data Closing Price, and Market Cap, and the moving averages 10 and 20 days have unit root which means that they are non-stationry. Good notes to have when comprising our model. Now that we have an understanding of what is stationry and non-stationary let's partition the data into training and testing out of time samples.

Before we do so let's do some more engineering and engineer our potential target variables. Create dummy ariables on when to buy and sel. First let's get a representation of the data to start.


```{r}
#names(btc_data)
#time_cols <- c("Date", "Bit_log_returns", "Bit_returns", "Bit_log_wtd_ret", "Bit_log_biwtd_ret",
               #"Bit_log_mtd_ret", "ema_log_ret_7", "ema_log_ret_14","ema_log_ret_30")

time_cols <- c("Date", "Bit_log_wtd_ret", "Bit_log_biwtd_ret",
               "Bit_log_mtd_ret")
btc_time <- btc_data[, (colnames(btc_data) %in% time_cols)]
btc_time <- melt(btc_time, id.vars = "Date")
btc_time <- btc_time[btc_time$Date > "2014-01-01" &    # Extract data frame subset
                   btc_time$Date < "2015-01-01", ]

time_line<- ggplot(btc_time,
                   aes(x = Date,
                       y = value,
                       col = variable,
                       alpha=0.25)) +
              geom_line() + ggtitle("Log WTD, BI-WTD, and MTD Returns")

time_cols <- c("Date", "ema_log_ret_7", "ema_log_ret_14","ema_log_ret_30")
btc_time <- btc_data[, (colnames(btc_data) %in% time_cols)]
btc_time <- melt(btc_time, id.vars = "Date")
btc_time <- btc_time[btc_time$Date > "2014-01-01" &    # Extract data frame subset
                   btc_time$Date < "2015-01-01", ]

time_line2<- ggplot(btc_time,
                   aes(x = Date,
                       y = value,
                       col = variable,
                       alpha=0.25)) +
              geom_line() + geom_hline(aes(yintercept = 0, colour = "Black")) + ggtitle("EMA 7, 14, and 30 days of Daily Log Returns")

# plot the simple returns as a density plot, box plot, and qqplot
wtd_dens <-  ggplot(btc_data, aes(x=Bit_log_wtd_ret)) +
              geom_histogram(aes(y = ..density..), bins= 20, colour = 1, fill= "blue", alpha = 0.6) +
              geom_density() + ggtitle("WTD Returns")

wtd_qq <- ggplot(btc_data, aes(sample = Bit_log_wtd_ret)) + stat_qq() + stat_qq_line() + ggtitle("WTD Returns")

wtd_box <- ggplot(btc_data, aes(y = Bit_log_wtd_ret)) + geom_boxplot(width = 0.5) + ggtitle("WTD Returns")

# plot the simple returns as a density plot, box plot, and qqplot
biwtd_dens <-  ggplot(btc_data, aes(x=Bit_log_biwtd_ret)) +
              geom_histogram(aes(y = ..density..), bins= 20, colour = 1, fill= "blue", alpha = 0.6) +
              geom_density() + ggtitle("Bi-WTD Returns")

biwtd_qq <- ggplot(btc_data, aes(sample = Bit_log_biwtd_ret)) + stat_qq() + stat_qq_line() + ggtitle("Bi-WTD Returns")

biwtd_box <- ggplot(btc_data, aes(y = Bit_log_biwtd_ret)) + geom_boxplot(width = 0.5) + ggtitle("Bi-WTD Returns")

# plot the simple returns as a density plot, box plot, and qqplot
mtd_dens <-  ggplot(btc_data, aes(x=Bit_log_mtd_ret)) +
              geom_histogram(aes(y = ..density..), bins= 20, colour = 1, fill= "blue", alpha = 0.6) +
              geom_density() + ggtitle("MTD Returns")

mtd_qq <- ggplot(btc_data, aes(sample = Bit_log_mtd_ret)) + stat_qq() + stat_qq_line() + ggtitle("MTD Returns")

mtd_box <- ggplot(btc_data, aes(y = Bit_log_mtd_ret)) + geom_boxplot(width = 0.5) + ggtitle("MTD Returns")

# plot the Moving average difference returns as a density plot, box plot, and qqplot
ma_diff_dens <-  ggplot(btc_data, aes(x=moving_avg_50_diff)) +
              geom_histogram(aes(y = ..density..), bins= 20, colour = 1, fill= "blue", alpha = 0.6) +
              geom_density() + ggtitle("50 Day MA 1 day differnence")

ma_diff_qq <- ggplot(btc_data, aes(sample = moving_avg_50_diff)) + stat_qq() + stat_qq_line() + ggtitle("50 Day MA 1 day differnence")

ma_diff_box <- ggplot(btc_data, aes(y = moving_avg_50_diff)) + geom_boxplot(width = 0.5) + ggtitle("50 Day MA 1 day differnence")

# plot the Moving Average as a density plot, box plot, and qqplot
ma7_diff_dens <-  ggplot(btc_data, aes(x=moving_avg_50_diff_7)) +
              geom_histogram(aes(y = ..density..), bins= 20, colour = 1, fill= "blue", alpha = 0.6) +
              geom_density() + ggtitle("50 Day MA 7 days differnence")

ma7_diff_qq <- ggplot(btc_data, aes(sample = moving_avg_50_diff_7)) + stat_qq() + stat_qq_line() + ggtitle("50 Day MA 7 days differnence")

ma7_diff_box <- ggplot(btc_data, aes(y = moving_avg_50_diff_7)) + geom_boxplot(width = 0.5) + ggtitle("50 Day MA 7 days differnence")

#Creat time series chart of the moving averages
time_cols <- c("Date", "moving_avg_50_diff_7", "moving_avg_50_diff")
btc_time <- btc_data[, (colnames(btc_data) %in% time_cols)]
btc_time <- melt(btc_time, id.vars = "Date")
btc_time <- btc_time[btc_time$Date > "2014-01-01" &    # Extract data frame subset
                   btc_time$Date < "2015-01-01", ]

time_line_ma<- ggplot(btc_time,
                   aes(x = Date,
                       y = value,
                       col = variable,
                       alpha=0.25)) +
              geom_line() + ggtitle("Moving Averages difference from Jan 2014 - Jan 2015")

grid.arrange(time_line, time_line2, ncol = 1, nrow = 2)
grid.arrange(wtd_dens, wtd_qq, wtd_box, biwtd_dens, biwtd_qq, biwtd_box, mtd_dens, mtd_qq, mtd_box, ncol = 3, nrow = 3)
grid.arrange(ma_diff_dens, ma_diff_qq, ma_diff_box, ma7_diff_dens, ma7_diff_qq, ma7_diff_box, ncol = 3, nrow = 2)
grid.arrange(time_line_ma, ncol = 1, nrow = 1)
```
Let's create a another ema set of features, this time using a bi-weekly cadence.

Additional Features:

ema_log_biret_7 - the exponential moving average using teh bi weekly returns. The time window for the value is at time t and t-7
ema_log_biret_14 - the exponential moving average using teh bi weekly returns. The time window for the value is at time t and t-14
ema_log_biret_30 - the exponential moving average using teh bi weekly returns. The time window for the value is at time t and t-30
```{r}
#Get the 7 day exponential moving average of daily returns
btc_data$ema_log_biret_7 <- EMA(btc_data$Bit_log_biwtd_ret, 7)
#Get the 14 day exponential moving average of daily returns
btc_data$ema_log_biret_14 <- EMA(btc_data$Bit_log_biwtd_ret, 14)
#Get the 30 day exponential moving average of daily returns
btc_data$ema_log_biret_30 <- EMA(btc_data$Bit_log_biwtd_ret, 30)

time_cols <- c("Date", "ema_log_biret_7", "ema_log_biret_14","ema_log_biret_30")
btc_time <- btc_data[, (colnames(btc_data) %in% time_cols)]
btc_time <- melt(btc_time, id.vars = "Date")
btc_time <- btc_time[btc_time$Date > "2014-01-01" &    # Extract data frame subset
                   btc_time$Date < "2015-01-01", ]

time_line3<- ggplot(btc_time,
                   aes(x = Date,
                       y = value,
                       col = variable,
                       alpha=0.25)) +
              geom_line() + geom_hline(aes(yintercept = 0, colour = "Black")) + ggtitle("EMA 7, 14, and 30 days of BI Weekly to date Log Returns")

time_cols <- c("Date", "Close")
btc_time <- btc_data[, (colnames(btc_data) %in% time_cols)]
btc_time <- btc_time[btc_time$Date > "2014-01-01" &    # Extract data frame subset
                   btc_time$Date < "2015-01-01", ]

time_line6<- ggplot(btc_time,
                   aes(x = Date,
                       y = Close,
                       alpha=0.25)) +
              geom_line() + ggtitle("Prices Jan 2014 - Jan 2015")

#plot the 50 day moving average
time_cols <- c("Date", "moving_avg_50")
btc_time <- btc_data[, (colnames(btc_data) %in% time_cols)]
btc_time <- btc_time[btc_time$Date > "2014-01-01" &    # Extract data frame subset
                   btc_time$Date < "2015-01-01", ]

time_line50<- ggplot(btc_time,
                   aes(x = Date,
                       y = moving_avg_50,
                       alpha=0.25)) +
              geom_line() + ggtitle("Prices Jan 2014 - Jan 2015")


ma_dens <-  ggplot(btc_data, aes(x=moving_avg_50)) +
              geom_histogram(aes(y = ..density..), bins= 20, colour = 1, fill= "blue", alpha = 0.6) +
              geom_density() + ggtitle("MTD Returns")

ma_qq <- ggplot(btc_data, aes(sample = moving_avg_50)) + stat_qq() + stat_qq_line() + ggtitle("MTD Returns")

ma_box <- ggplot(btc_data, aes(y = moving_avg_50)) + geom_boxplot(width = 0.5) + ggtitle("MTD Returns")

grid.arrange(time_line2, time_line3, time_line6, time_line50, ncol = 1, nrow = 4)
grid.arrange(ma_dens, ma_qq, ma_box, ncol = 3, nrow = 1)
```
Get the ema for monthly returns, plot it against the ema for Bi wtd and clossing prices of Bitcoin
```{r}
#Get the 7 day exponential moving average of daily returns
btc_data$ema_log_mtdret_7 <- EMA(btc_data$Bit_log_mtd_ret, 7)
#Get the 14 day exponential moving average of daily returns
btc_data$ema_log_mtdret_14 <- EMA(btc_data$Bit_log_mtd_ret, 14)
#Get the 30 day exponential moving average of daily returns
btc_data$ema_log_mtdret_30 <- EMA(btc_data$Bit_log_mtd_ret, 30)

time_cols <- c("Date", "ema_log_mtdret_7", "ema_log_mtdret_14","ema_log_mtdret_30")
btc_time <- btc_data[, (colnames(btc_data) %in% time_cols)]
btc_time <- melt(btc_time, id.vars = "Date")
btc_time <- btc_time[btc_time$Date > "2014-01-01" &    # Extract data frame subset
                   btc_time$Date < "2015-01-01", ]

time_line5<- ggplot(btc_time,
                   aes(x = Date,
                       y = value,
                       col = variable,
                       alpha=0.25)) +
              geom_line() + geom_hline(aes(yintercept = 0, colour = "Black")) + ggtitle("EMA 7, 14, and 30 days of MTD Log Returns")

time_cols <- c("Date", "Close")
btc_time <- btc_data[, (colnames(btc_data) %in% time_cols)]
btc_time <- btc_time[btc_time$Date > "2014-01-01" &    # Extract data frame subset
                   btc_time$Date < "2015-01-01", ]


grid.arrange(time_line3, time_line5, time_line6, ncol = 1, nrow = 3)
```
Get the graphs of the EMA values that were recently created.
```{r}
# plot the simple returns as a density plot, box plot, and qqplot
ema_mtd_dens <-  ggplot(btc_data, aes(x=ema_log_mtdret_30)) +
              geom_histogram(aes(y = ..density..), bins= 20, colour = 1, fill= "blue", alpha = 0.6) +
              geom_density() + ggtitle("EMA 30 MTD Returns")

ema_mtd_qq <- ggplot(btc_data, aes(sample = ema_log_mtdret_30)) + stat_qq() + stat_qq_line() + ggtitle("EMA 30 MTD Returns")

ema_mtd_box <- ggplot(btc_data, aes(y = ema_log_mtdret_30)) + geom_boxplot(width = 0.5) + ggtitle("EMA 30 MTD Returns")

# plot the simple returns as a density plot, box plot, and qqplot
ema_biwtd_dens <-  ggplot(btc_data, aes(x=ema_log_biret_30)) +
              geom_histogram(aes(y = ..density..), bins= 20, colour = 1, fill= "blue", alpha = 0.6) +
              geom_density() + ggtitle("EMA BI-WTD Returns")

ema_biwtd_qq <- ggplot(btc_data, aes(sample = ema_log_biret_30)) + stat_qq() + stat_qq_line() + ggtitle("EMA 30 BI-WTD Returns")

ema_biwtd_box <- ggplot(btc_data, aes(y = ema_log_biret_30)) + geom_boxplot(width = 0.5) + ggtitle("EMA 30 BI-WTD Returns")

grid.arrange(ema_mtd_dens, ema_mtd_qq, ema_mtd_box, ema_biwtd_dens, ema_biwtd_qq, ema_biwtd_box, ncol = 3, nrow = 2)
```
Shift the week to date, bi week to date, and month to date up the number of units it used to calclute the metric. (i.e week date used t and t-7 to calculate. Shift the rows up replacing the previous null values). This will now give the future returns at time t.
```{r}
#Shift the columns back n days so current features align with the labeled future values
n <- nrow(btc_data) 
btc_data$Bit_log_wtd_ret_lag <- c(btc_data$Bit_log_wtd_ret[8:n], rep(NA,7))
btc_data$Bit_log_biwtd_ret_lag <- c(btc_data$Bit_log_biwtd_ret[15:n], rep(NA,14))
btc_data$Bit_log_mtd_ret_lag <- c(btc_data$Bit_log_mtd_ret[31:n], rep(NA,30))
#Create the indicators on when to buy when to sell and when to hold
btc_data$buy_sell_ind_wtd <- ifelse(btc_data$Bit_log_wtd_ret_lag >=  0.2, 1, ifelse(btc_data$Bit_log_wtd_ret_lag <=  -0.15, -1, 0))
btc_data$buy_sell_ind_biwtd <- ifelse(btc_data$Bit_log_biwtd_ret_lag >=  0.2, 1, ifelse(btc_data$Bit_log_biwtd_ret_lag <=  -0.15, -1, 0))
btc_data$buy_sell_ind_mtd <- ifelse(btc_data$Bit_log_mtd_ret_lag >=  0.2, 1, ifelse(btc_data$Bit_log_mtd_ret_lag <=  -0.15, -1, 0))

# Make the new columns as factor type
btc_data$buy_sell_ind_wtd <- as.factor(btc_data$buy_sell_ind_wtd)
btc_data$buy_sell_ind_biwtd <- as.factor(btc_data$buy_sell_ind_biwtd)
btc_data$buy_sell_ind_mtd <- as.factor(btc_data$buy_sell_ind_mtd)
# remove the rows with null values at the beginning and end of the dataframe.
btc <- btc_data[complete.cases(btc_data),]
#Partition the data into training and testing sets.
train_indx <- round(0.7 * nrow(btc))
btc_train <- btc[1:train_indx,]
test_indx <- train_indx + 1
n_btc <- nrow(btc)
btc_test <- btc[test_indx:n_btc,]
```

Look at the relationships between EMA Values and their expected future returns
```{r}
day <- ggplot(btc_train, aes(x=Bit_log_mtd_ret, y=Bit_log_wtd_ret_lag, color=buy_sell_ind_wtd)) + 
        geom_point(alpha=0.5, shape=18) + geom_vline(aes(xintercept = 0)) + geom_hline(aes(yintercept = 0))

week <- ggplot(btc_train, aes(x=ema_log_biret_30, y=Bit_log_wtd_ret_lag, color=buy_sell_ind_wtd)) + 
        geom_point(alpha=0.5, shape=18) + geom_vline(aes(xintercept = 0)) + geom_hline(aes(yintercept = 0))

month <- ggplot(btc_train, aes(x=Bit_log_biwtd_ret_lag, y=Bit_log_mtd_ret_lag, color=buy_sell_ind_mtd)) + 
        geom_point(alpha=0.5, shape=18) + geom_vline(aes(xintercept = 0)) + geom_hline(aes(yintercept = 0))

grid.arrange(day, week, month, ncol = 1, nrow = 3)

names(btc_train)
cor_feat <- cor(btc_train[, 5:34])
corrplot(cor_feat, method="color", addCoef.col = 0.5, number.cex = 0.5, tl.cex = 0.5)
```
plot the buy and sell points on a chart of prices.
```{r}
time_cols <- c("Date", "Close", "buy_sell_ind_mtd")
btc_time <- btc_data[, (colnames(btc_data) %in% time_cols)]
btc_time <- btc_time[btc_time$Date > "2013-06-01" &    # Extract data frame subset
                   btc_time$Date < "2015-01-01", ]
btc_points <- btc_time[btc_time$buy_sell_ind_mtd == -1, ]
plot(x = btc_time$Date,
    y = btc_time$Close, main = "Sell Indicators between Jun 2013 - Jan 2015")
points(x = btc_points$Date, y=btc_points$Close, col="red")
```
```{r}
time_cols <- c("Date", "Close", "buy_sell_ind_mtd")
btc_time <- btc_data[, (colnames(btc_data) %in% time_cols)]
btc_time <- btc_time[btc_time$Date > "2013-06-01" &    # Extract data frame subset
                   btc_time$Date < "2015-01-01", ]
btc_points <- btc_time[btc_time$buy_sell_ind_mtd == 1, ]
plot(x = btc_time$Date,
    y = btc_time$Close, main = "Buy Indicators between Jun 2013 - Jan 2015")
points(x = btc_points$Date, y=btc_points$Close, col="green")
```
After some research we will use the xgboost algorithim to classiy when to buy and when to sell based where we think positive returns will be a month out from the point in time.

XGBoost can be used for either regression or classification problems. In our case we engineered our target variable with 0s and 1s, making it a classification problem. We will see if using xgboost as classification model would be able to spot good buy moments in the data.

Xgboost in time series: https://machinelearningmastery.com/xgboost-for-time-series-forecasting/

Why model using log returns: https://medium.datadriveninvestor.com/why-we-use-log-returns-for-stock-returns-820cec4510ba
# Main
```{r}
str(btc_train)
rm_cols <- c("SNo", "Name", "Symbol", "High", "Low", "Open", "Close", "Marketcap", "buy_sell_ind_wtd", "buy_sell_ind_biwtd",
             "Bit_log_wtd_ret_lag", "Bit_log_biwtd_ret_lag", "Bit_log_mtd_ret_lag", "buy_sell_ind_mtd", "moving_avg_10",
             "moving_avg_20")
#Predict on when is a good time to buy
btc_train$buy_indicator <- ifelse(btc_train$buy_sell_ind_mtd == 1, 1, 0)
btc_test$buy_indicator <- ifelse(btc_test$buy_sell_ind_mtd == 1, 1, 0)
btc_train <- btc_train[, !(colnames(btc_train) %in% rm_cols)]
btc_test <- btc_test[, !(colnames(btc_test) %in% rm_cols)]

# Convert the date into integers of month day and year, each with its own column to preserve the temporal aspect of the data

btc_train <- btc_train %>%
    dplyr::mutate(., 
                  months = lubridate::month(Date),
                  years = lubridate::year(Date),
                  days = lubridate::day(Date))

btc_test <- btc_test %>%
    dplyr::mutate(., 
                  months = lubridate::month(Date),
                  years = lubridate::year(Date),
                  days = lubridate::day(Date))
#Convert the days column to numeric
btc_test$days <- as.numeric(btc_test$days)
btc_train$days <- as.numeric(btc_train$days)
btc_test$months <- as.numeric(btc_test$months)
btc_train$months <- as.numeric(btc_train$months)
btc_test$year <- as.numeric(btc_test$years)
btc_train$years <- as.numeric(btc_train$years)
#create the matrix values for the xgboost
x_train <- btc_train %>%
                    dplyr::select(months, days, years, Volume, Bit_log_returns, Bit_returns, Bit_days_range, Bit_log_wtd_ret, Bit_log_biwtd_ret,
                    Bit_log_mtd_ret, ema_log_ret_7, ema_log_ret_14, ema_log_ret_30, Bit_rsi_3, Bit_rsi_60, Bit_rsi_90,
                    ema_log_biret_7, ema_log_biret_14, ema_log_biret_30, ema_log_mtdret_7, ema_log_mtdret_14, ema_log_mtdret_30, moving_avg_50,
                    moving_avg_50_diff, moving_avg_50_diff_7) %>%
                  as.matrix()

x_test <- btc_test %>% 
                  dplyr::select(months, days, years, Volume, Bit_log_returns, Bit_returns, Bit_days_range, Bit_log_wtd_ret, Bit_log_biwtd_ret,
                  Bit_log_mtd_ret, ema_log_ret_7, ema_log_ret_14, ema_log_ret_30, Bit_rsi_3, Bit_rsi_60, Bit_rsi_90,
                  ema_log_biret_7, ema_log_biret_14, ema_log_biret_30, ema_log_mtdret_7, ema_log_mtdret_14, ema_log_mtdret_30, moving_avg_50,
                  moving_avg_50_diff, moving_avg_50_diff_7) %>%
                as.matrix()
y_train <- as.factor(btc_train$buy_indicator)
y_test <- as.factor(btc_test$buy_indicator)
```
link:
https://www.r-bloggers.com/2019/09/time-series-forecasting-with-random-forest/
```{r}
#nrow(btc_train)
set.seed(123)
xgb_trcontrol <- caret::trainControl(
   method = "timeslice",
   initialWindow = 200,
   horizon = 70,
   fixedWindow = FALSE, 
   allowParallel = FALSE
)
 
xgb_grid <- base::expand.grid(
   list(
    nrounds = c(150, 250),
    max_depth = c(20, 50, 100), # maximum depth of a tree
    colsample_bytree = seq(0.5), # subsample ratio of columns when construction each tree
    eta = 0.1, # learning rate
    gamma = 0, # minimum loss reduction
    min_child_weight = 1,  # minimum sum of instance weight (hessian) needed ina child
    subsample = 1 # subsample ratio of the training instances
))
```
```{r, warning=FALSE}
xgb_model <- caret::train(
   x_train, y_train,
   trControl = xgb_trcontrol,
   tuneGrid = xgb_grid,
   method = "xgbTree",
   nthread = 1,
   objective = "binary:logistic"
)
```
```{r}
xgb_model$bestTune
```
```{r}
importance_matrix <- xgb.importance(model = xgb_model$finalModel)
print(importance_matrix)
xgb.plot.importance(importance_matrix = importance_matrix)
```

As we can see from the feature
## Methodology
```{r}
train_predicted <- xgb_model %>% predict(x_train)
print("The accuracy score for the buy indicator is bellow:")
Accuracy(train_predicted, btc_train$buy_indicator)

#Confussion matrix
confMat1 <- ConfusionMatrix(train_predicted, btc_train$buy_indicator)
confMat1

```
With results like that our model could be over fitted.
```{r}
test_predicted <- xgb_model %>% predict(x_test)
print("The accuracy score for the buy indicator is bellow:")
Accuracy(test_predicted, btc_test$buy_indicator)

#Confussion matrix
confMat2 <- ConfusionMatrix(test_predicted, btc_test$buy_indicator)
confMat2

roc_offer <- roc(test_predicted, as.integer(btc_test$buy_indicator))
plot(roc_offer)

btc_test$pred1 <- test_predicted

time_cols <- c("Date", "Close")
btc_sub <- btc_data[, (colnames(btc_data) %in% time_cols)]
btc_test <- left_join(btc_test, btc_sub, by="Date")
time_cols <- c("Date", "Close", "pred1")
btc_time <- btc_test[, (colnames(btc_test) %in% time_cols)]
btc_points <- btc_time[btc_time$pred1 == 1, ]
plot(x = btc_time$Date,
    y = btc_time$Close, main = "Buy Indicators between Jun 2013 - Jan 2015")
points(x = btc_points$Date, y=btc_points$Close, col="green")



```
As we can see our model did perfectly well at segmenting the data perfectly. However, it did poorly against the testing data. The columns that we care about most is the ratio between true positve false positives from a business standpoint. Reason being is that, those 2 counts indicate a buy order was made. From a business standpoint we would want to first ensure that a buy indicator indicates a winning buy order, being that the returns will be positive.

This time lets train the model stripping some of the lesser important features, which should slim our data down. Then we will increase the window size of the testing in the cross validation step.
```{r}
#create the matrix values for the xgboost
x_train_2 <- btc_train %>%
                    dplyr::select(months, days, years,
                    Bit_log_mtd_ret, ema_log_ret_14, Bit_rsi_90,
                    ema_log_mtdret_30, moving_avg_50_diff_7) %>%
                  as.matrix()

x_test_2 <- btc_test %>% 
                  dplyr::select(months, days, years,
                    Bit_log_mtd_ret, ema_log_ret_14, Bit_rsi_90,
                    ema_log_mtdret_30, moving_avg_50_diff_7) %>%
                as.matrix()

set.seed(123)
xgb_trcontrol_2 <- caret::trainControl(
   method = "timeslice",
   initialWindow = 700,
   horizon = 200,
   fixedWindow = FALSE, 
   allowParallel = FALSE
)
 
xgb_grid_2 <- base::expand.grid(
   list(
    nrounds = c(100, 200),
    max_depth = c(10, 15, 20), # maximum depth of a tree
    colsample_bytree = seq(0.5), # subsample ratio of columns when construction each tree
    eta = 0.1, # learning rate
    gamma = 0, # minimum loss reduction
    min_child_weight = 1,  # minimum sum of instance weight (hessian) needed ina child
    subsample = 1 # subsample ratio of the training instances
))
```



```{r, warning=FALSE}
xgb_model_2 <- suppressMessages(caret::train(
   x_train, y_train,
   trControl = xgb_trcontrol_2,
   tuneGrid = xgb_grid_2,
   method = "xgbTree",
   nthread = 1,
   objective = "binary:logistic"
))
```
get the results.

```{r}
train_predicted_2 <- xgb_model_2 %>% predict(x_train)
print("The accuracy score for the buy indicator is bellow:")
Accuracy(train_predicted_2, btc_train$buy_indicator)

#Confussion matrix
confMat1_2 <- ConfusionMatrix(train_predicted_2, btc_train$buy_indicator)
confMat1_2


test_predicted_2 <- xgb_model_2 %>% predict(x_test)
print("The accuracy score for the buy indicator is bellow:")
Accuracy(test_predicted_2, btc_test$buy_indicator)

#Confussion matrix
confMat2_2 <- ConfusionMatrix(test_predicted_2, btc_test$buy_indicator)
confMat2_2

```
Check the kind of trade that the algorithim predicted.
```{r}
time_cols <- c("Date", "Close")
btc_sub <- btc_data[, (colnames(btc_data) %in% time_cols)]
btc_test <- left_join(btc_test, btc_sub, by="Date")
btc_test$pred_1 <- test_predicted_2
time_cols <- c("Date", "Close", "pred_1")
btc_time <- btc_test[, (colnames(btc_test) %in% time_cols)]
btc_points <- btc_time[btc_time$pred_1 == 1, ]
plot(x = btc_time$Date,
    y = btc_time$Close, main = "Buy Indicators between Jun 2013 - Jan 2015")
points(x = btc_points$Date, y=btc_points$Close, col="green")

```

```{r}
max(btc_test$Date)
```

### Initial Exploration
#### Assumptions

#### Removing Outliers

#### Results and Conclusions

# Challenges and Limitations

# Further Discussion and Research

# References
